### 2024-12-06 manuscript.tex
Addressing comments

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is a 5-km version of the Icosahedral Nonhydrostatic Weather and
Climate Model (ICON). The high resolution allows for parameterizations of
convection and clouds to be avoided, relying instead on explicitly resolved
processes, except for, among others, microphysics and turbulence.  Standard-resolution models have substantial cloud biases over the
Southern Ocean (SO), affecting radiation and sea surface temperature.  We
evaluated SO clouds in ICON and the ERA5 and MERRA-2 reanalyses.  The SO is
dominated by low clouds, which cannot be observed accurately from space due to
overlapping clouds, attenuation, and ground clutter.  Instead, we analysed
about 2400 days of lidar observations from 31 voyages and a station during
2010--2021 using a ground-based lidar simulator.  ICON and the reanalyses
underestimate the total cloud fraction by about 10 and 20\%, respectively. ICON
and ERA5 overestimate the cloud occurrence peak at about 500 m, potentially
explained by their lifting condensation levels being too high.  The reanalyses
strongly underestimate near-surface clouds or fog.  MERRA-2 tends to
underestimate cloud occurrence at all heights. Conditions of weak stability are associated with the largest errors in ICON and the reanalyses.  In daily cloud cover, ICON and
the reanalyses tend to be about 1 and 2 oktas clearer, respectively. Compared
to radiosondes, potential temperature is accurate in the reanalyses, but ICON
underestimates stability over the low-latitude SO and too humid in the boundary
layer. MERRA-2 is too humid at all heights. SO cloud biases remain a
substantial issue in the GSRM, but are an improvement over the lower-resolution
reanalyses. Explicitly resolved convection and cloud processes (except microphysics) were not enough
to address the model cloud biases.

@@

The Southern Ocean (SO) is known to be a problematic region for climate model
biases due to a lack of surface and in situ observations and being a lower
priority region for numerical weather prediction (NWP) and climate model
development because of its distance from populated areas.  Nevertheless,
radiation biases and changes over an area of its size have a substantial
influence on the global climate, and the SO is an important part of the global
ocean conveyor belt. In general, marine clouds have a disproportionate effect on top of
atmosphere (TOA) SW radiation due to the relatively low albedo of the sea
surface.  The relative longitudinal symmetry of the SO means that model cloud
biases tend to be similar across longitudes.  Here, we conventionally refer to
the SO as ocean regions south of 40°S, low-latitude SO as 40--55°S and
high-latitude SO as south of 55°S.

@@

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/lts_dist.pdf}
\caption{
Lower tropospheric stability (LTS) distribution in \textbf{(a)} ERA5 and
\textbf{(b)} MERRA-2 calculated for the 31 voyage tracks and one station from
the highest instantaneous temporal resolution data available. Shown is also the
chosen dividing threshold of 12 K for relatively stable and unstable
conditions.
}
\label{fig:lts}
\end{figure}

@@

We partitioned our data into two mutually exclusive subsets by cyclonic
activity. For this purpose, we used a cyclone tracking algorithm to identify
extratropical and polar cyclones (ECs and PCs) over the SO in the reanalysis
and ICON data. We used the open source cyclone tracking package CyTRACK
\citep{perez-alarcon2024}.  Generally, what constitutes an EC is considered
relatively arbitrary due to the very large variability of ECs \citep{neu2013}.
We used the mean sea level pressure field and horizontal wind speed fields as
input to the CyTRACK algorithm. The algorithm uses pressure and wind speed
thresholds as well as tracking across time steps to identify cyclone centres
and radii in each time step. With this information, we could classify geographical areas as
either cyclonic or non-cyclonic. Due to a relatively small total area covered by cyclones (as identified by the cyclone centre and radius),
we chose a circle of a double radius (relative to one identified by CyTRACK)
centred at the cyclone centre as a cyclonic area for every time step and
cyclone. All other areas were identified as non-cyclonic. For identifying
cyclones in the observations and the reanalyses, ERA5 pressure and wind fields
were used as the input to CyTRACK.  This is justified by the fact that the
large-scale pressure and wind fields in ERA5 are likely sufficiently close to
reality. For identifying cyclones in ICON, its own pressure and wind fields
were used as the input to CyTRACK, because the model is free-running, and thus
the pressure and wind fields are different from reality.

In addition to the above, we partitioned our data into two mutually exclusive
subsets based on LTS, which is derived as the difference between the potential temperature at 700 hPa
and the surface.  Based on a histogram of LTS in ERA5 and MERRA-2 calculated at
all voyage tracks and stations (Fig.  \ref{fig:lts}), we determined a dividing
threshold of 12 K for relatively unstable (< 12 K) and relatively stable (>= 12
K) conditions.

@@

\begin{figure}[p!]
\centering
\includegraphics[width=\textwidth]{img/cyc_stab_dist.pdf}
\caption{
Geographical distribution of \textbf{(a, b)} cyclonic days and \textbf{(b, d)}
relatively stable (LTS > 12 K) time steps in \textbf{(a, c)} ERA5 in years
2010--2013 (inclusive) and \textbf{(b, d)} ICON in model years 2021--2023 (free
running). Cyclonic days are expressed as a fraction of the number of days with
cyclonic activity, defined as grid points located within a double radius of any
cyclone on a given day (UTC), as identified by CyTRACK.
}
\label{fig:cyclone-stability}
\end{figure}

@@

Fig. \ref{fig:cyclone-stability}c, d show a geographical distribution of the
relatively stable and unstable conditions as determined by the LTS (Section
\ref{sec:cyclone-stability}). Relatively unstable conditions are prevalent
in the mid-to-high SO (with respect to our SO partitioning; 50--65°S), which might be explained by the relatively cold
near-surface air overlying the relatively warm sea surface. Relatively
stable conditions are prevalent elsewhere over the SO. The distribution is also
less zonally symmetric than the cyclonic activity.  In the high-latitude SO,
the presence of sea ice might have substantial stabilising effect
\citep{knight2024}. The ERA5 reanalysis is also substantially more stable than
ICON across the whole region.

\begin{figure}[p!]
\centerline{
\includegraphics[width=1.06\textwidth]{img/cloud_occurrence_panel.pdf}
}
\caption{
Cloud occurrence by height for the 31 voyages and one sub-Antarctic
station (MICRE) in observations (O) and simulated by the ALCF from the ICON
model (I), MERRA‐2 (M), and ERA5 reanalysis data (E). The numbers in the legend
indicate the total cloud fraction and the number of days of data.
}
\label{fig:cloud-occurrence-panel}
\end{figure}

@@

For all observations considered (Fig. \ref{fig:cloud-occurrence}a), the data show
cloud occurrence peaking at the surface, whereas the models show a higher peak (at
about 500 m). The models generally underestimate the total cloud fraction by 10--20\% and show a strong reduction in cloud occurrence near the surface, which is not
supported by the observations. ICON and ERA5 overestimate cloud occurrence at
their peak (between 0--1 km). Above 1 km, ICON and MERRA-2 underestimate cloud
occurrence, but ERA5 is very accurate. The exaggerated peak in models is partly
supported by the lifting condensation level (LCL) distribution, which peaks
higher in the models than in the observations (at the surface), although this
is not very pronounced.

When subsetted by latitude (Fig.
\ref{fig:cloud-occurrence}b, c), we see that the low-latitude SO zone displays
a stronger peak of cloud occurrence near the surface than the high-latitude SO
zone, and this could be because higher latitudes have more unstable atmospheric
profiles. The low- and high-latitude SO zones show similar biases in models as
in the general case, but ERA5 does not overestimate the peak in the
low-latitude SO zone (near-surface cloud occurrence is still strongly
underestimated).

@@

When subsetted by relatively stable and unstable conditions (Fig.
\ref{fig:cloud-occurrence}f, g), as defined in Section
\ref{sec:cyclone-stability}, we see that in relatively stable
situations cloud occurrence peaks strongly at the surface in observations,
compared to relatively unstable situations, where the peak is more diffuse
between 0 and 1 km.  In relatively stable
situations, the models do a fairly good job, but overestimate cloud
occurrence at the peak below 1 km; above 1 km, they show similar biases as in
the general case.  In relatively unstable situations, the bias in ICON is
very pronounced, with a much stronger peak at about 500 m, ERA5 underestimates cloud occurrence below 1 km (especially near the surface), and
MERRA-2 underestimates cloud occurrence even more strongly.

In all situations, even when the models overestimate cloud occurrence at some
altitudes, they always substantially underestimate the total cloud fraction.
ICON can be generally characterised as substantially overestimating cloud
occurrence below 1 km and underestimating above, underestimating the total
cloud fraction, and showing greatest biases in relatively unstable and
non-cyclonic conditions. It also shows a peak of cloud occurrence at higher
altitude than observations (500 m vs. near the surface), and correspondingly,
its LCL tends to be higher. MERRA-2 can be generally characterised as
underestimating cloud occurrence at nearly all altitudes as well as the total
cloud fraction, but mostly above and below 500 m (the peak at 500 m is
well-represented). MERRA-2 displays the largest errors relative to observations in the low-latitude SO zone and in relatively unstable situations. ERA5 can be generally characterised as
representing cloud occurrence correctly above about 1--1.5 km, overestimating
below, but underestimating near-surface cloud occurrence (0--500 m). The total
cloud fraction is strongly underestimated in all situations. ERA5 has a tendency
towards underestimation in the low-latitude SO zone and relatively unstable
situations; conversely, overestimating in the high-latitude SO zone and relatively stable conditions.

\begin{figure}[p!]
\centering
\includegraphics[width=\textwidth]{img/cl_agg.pdf}
\caption{
Cloud occurrence by height calculated as the average of all voyages and
stations for the observed (OBS) and simulated lidar cloud mask, and lifting
condensation level (LCL) distribution. The total cloud fraction (CF), average
shortwave (SW), and longwave (LW) are shown in the legend, and the relative
frequency of occurrence (RFO) of the subset is shown below.  The bands span the
16$^\mathrm{th}$--84$^\mathrm{th}$ percentile calculated from the set of all
voyages and stations. The subsets \textbf{(d--g)} are defined in Section
\ref{sec:cyclone-stability}.
}
\label{fig:cloud-occurrence}
\end{figure}

@@

When analysed by subsets, observations in the cyclonic subset show the highest
cloud cover, with 8 oktas occurring on one half of such days (Fig.
\ref{fig:cloud-cover}d).  This sensitivity to cyclonic conditions is not observed in ICON or the reanalyses. Interestingly, clear sky days (0 oktas) also have a local maximum peaking
at about 15\% in this subset.  When we contrast the low- and high-latitude
zones, we see that the high-latitude zone tends to have greater cloud cover,
peaking at 8 oktas (Fig.  \ref{fig:cloud-cover}c). The high-latitude zone also
has almost no clear sky or small cloud cover cases (0--4 oktas). ICON and the
reanalyses represent this characteristic of the distribution well for
0--3 oktas, but otherwise show biases similar to the general case. One of the
greatest biases is present in ERA5 in the relatively unstable subset, in which
ERA5 peaks at 3 oktas, while the observations peak at 7 oktas and show
negligible cloud cover below 5 oktas.

@@

Relative humidity displays much larger biases. In all subsets, ICON is too
humid in the first 1 km by about 5\% but very accurate above, except for the 40--55°S zone
and unstable conditions (Fig. \ref{fig:potential-temperature}b, g),  where it
is too dry between about 1 and 3 km. MERRA-2, on the other hand, is much more
humid than observations at all altitudes and in all subsets, by up to about
20\% at 5 km.

@@

Compared to lidar observations, the daily cloud cover tends to be about 1 okta
lower in ICON and 2 oktas lower in the reanalyses.  Unstable conditions are
associated with some of the greatest biases, especially in ERA5.  The models
also underestimate the cloud cover very strongly in cyclonic conditions, which
are very cloudy in the observations (8 oktas), but much less so in the models.

>>>>

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is a 5-km version of the Icosahedral Nonhydrostatic Weather and
Climate Model (ICON). The high resolution means that parameterizations of
convection and clouds including subgrid-scale clouds are forgone, relying on
explicit their simulation, but still using microphysics and turbulence
parametrization.  Standard-resolution (10--100-km) models, which use
subgrid-scale convection and cloud parametrization, have substantial cloud
biases over the Southern Ocean (SO), affecting radiation and sea surface
temperature.  We evaluated SO clouds in ICON and the ERA5 and MERRA-2
reanalyses.  The SO is dominated by low clouds, which cannot be observed
accurately from space due to overlapping clouds, attenuation, and ground
clutter.  Instead, we analysed about 2400 days of lidar observations from 31
voyages and a station during 2010--2021 using a ground-based lidar simulator.
ICON and the reanalyses underestimate the total cloud fraction by about 10 and
20\%, respectively. ICON and ERA5 overestimate the cloud occurrence peak at
about 500 m, potentially explained by their lifting condensation levels being
too high.  The reanalyses strongly underestimate near-surface clouds or fog.
MERRA-2 tends to underestimate cloud occurrence at all heights.  In daily cloud
cover, ICON and the reanalyses tend to be about 1 and 2 oktas clearer,
respectively.  Conditions of weak stability are associated with the largest
errors in ICON and the reanalyses.  Compared to radiosondes, potential
temperature is accurate in the reanalyses, but ICON underestimates stability
over the low-latitude SO and too humid in the boundary layer. MERRA-2 is too
humid at all heights. SO cloud biases remain a substantial issue in the GSRM,
but are an improvement over the lower-resolution reanalyses. Explicitly
resolved convection and cloud processes (except microphysics) were not enough
to address the model cloud biases.

@@

The Southern Ocean (SO) is known to be a problematic region for climate model
biases due to a lack of surface and in situ observations and being a lower
priority region for numerical weather prediction (NWP) and climate model
development because of its distance from populated areas.  Nevertheless,
radiation biases and changes over an area of its size have a substantial
influence on the global climate, such as affecting the Earth radiation balance
and carbon cycle, and the SO is an important part of the global ocean conveyor
belt. In general, marine clouds have a disproportionate effect on top of
atmosphere (TOA) SW radiation due to the relatively low albedo of the sea
surface.  The relative longitudinal symmetry of the SO means that model cloud
biases tend to be similar across longitudes.  Here, we conventionally refer to
the SO as ocean regions south of 40°S, low-latitude SO as 40--55°S and
high-latitude SO as south of 55°S.

@@

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/lts_dist.pdf}
\caption{
Lower tropospheric stability (LTS) distribution in \textbf{(a)} ERA5 and
\textbf{(b)} MERRA-2 calculated for the 31 voyage tracks and one station from
the highest instantaneous temporal resolution data available. Shown is also the
chosen dividing threshold of 12 K for conditions weak and strong stability.
}
\label{fig:lts}
\end{figure}

@@

We partitioned our data into two mutually exclusive subsets by cyclonic
activity. For this purpose, we used a cyclone tracking algorithm to identify
extratropical and polar cyclones (ECs and PCs) over the SO in the reanalysis
and ICON data. We used the open source cyclone tracking package CyTRACK
\citep{perez-alarcon2024}.  Generally, what constitutes an EC is considered
relatively arbitrary due to the very large variability of ECs \citep{neu2013}.
We used the mean sea level pressure field and horizontal wind speed fields as
input to the CyTRACK algorithm. The algorithm uses pressure and wind speed
thresholds as well as tracking across time steps to identify cyclone centres
and radii in each time step. With this information, we could classify
geographical areas as either cyclonic or non-cyclonic. Due to a relatively
small total area covered by cyclones (as identified by the cyclone centre and
radius), we chose a circle of a double radius (relative to one identified by
CyTRACK) centred at the cyclone centre as a cyclonic area for every time step
and cyclone. All other areas were identified as non-cyclonic. For identifying
cyclones in the observations and the reanalyses, ERA5 pressure and wind fields
were used as the input to CyTRACK.  This is justified by the fact that the
large-scale pressure and wind fields in ERA5 are likely sufficiently close to
reality. For identifying cyclones in ICON, its own pressure and wind fields
were used as the input to CyTRACK, because the model is free-running, and thus
the pressure and wind fields are different from reality. Subsetting by
proximity to cyclones is a relatively crude measure because it does not take
into account the different sectors of cyclones, which are commonly associated
with different weather situations. However, this was a choice made for
simplicity of the analysis, given the amount of data.

In addition to the above, we partitioned our data into two mutually exclusive
subsets based on LTS, which is derived as the difference between the potential
temperature at 700 hPa and the surface.  Based on a histogram of LTS in ERA5
and MERRA-2 calculated at all voyage tracks and stations (Fig.  \ref{fig:lts}),
we determined a statistically-based dividing threshold of 12 K for weak
stability (< 12 K) and strong stability (>= 12 K) conditions.

@@

\begin{figure}[p!]
\centering
\includegraphics[width=\textwidth]{img/cyc_stab_dist.pdf}
\caption{
Geographical distribution of \textbf{(a, b)} cyclonic days and \textbf{(b, d)}
strong stability (LTS > 12 K) time steps in \textbf{(a, c)} ERA5 in years
2010--2013 (inclusive) and \textbf{(b, d)} ICON in model years 2021--2023 (free
running). Cyclonic days are expressed as a fraction of the number of days with
cyclonic activity, defined as grid points located within a double radius of any
cyclone on a given day (UTC), as identified by CyTRACK.
}
\label{fig:cyclone-stability}
\end{figure}

@@

Fig. \ref{fig:cyclone-stability}c, d show a geographical distribution of the
conditions of weak and strong stability as determined by the LTS (Section
\ref{sec:cyclone-stability}). Conditions of weak stability are prevalent in the
mid-to-high SO (with respect to our SO partitioning; 50--65°S), which might be
explained by the relatively cold near-surface air overlying the relatively warm
sea surface. Conditions of strong stability are prevalent elsewhere over the
SO. The distribution is also less zonally symmetric than the cyclonic activity.
In the high-latitude SO, the presence of sea ice might have substantial
stabilising effect \citep{knight2024}. The ERA5 reanalysis is also
substantially more stable than ICON across the whole region.

\begin{figure}[p!]
\centerline{
\includegraphics[width=1.06\textwidth]{img/cloud_occurrence_panel.pdf}
}
\caption{
Cloud occurrence by height for the 31 voyages and one sub-Antarctic
station (MICRE) in observations (O) and simulated by the ALCF from the ICON
model (I), MERRA‐2 (M), and ERA5 reanalysis data (E). The numbers in the legend
indicate the total cloud fraction and the number of days of data. Multiple lines
of ICON profiles are for each of the four years of model data available.
}
\label{fig:cloud-occurrence-panel}
\end{figure}

@@

For all observations considered (Fig. \ref{fig:cloud-occurrence}a), the data
show cloud occurrence peaking at the surface, whereas the models show a higher
peak (at about 500 m). The models generally underestimate the total cloud
fraction by 10--20\% and show a strong reduction in cloud occurrence near the
surface, which is not supported by the observations. ICON and ERA5 overestimate
cloud occurrence at their peak (between 0--1 km). Above 1 km, ICON and MERRA-2
underestimate cloud occurrence, but ERA5 is very accurate. The exaggerated peak
in models is partly supported by the lifting condensation level (LCL)
distribution, which peaks higher in the models than in the observations (at the
surface), although this is not very pronounced. This is indicative of
near-surface relatively humidity being often close to saturation in the
observations but not in the models.

When subsetted by latitude (Fig.  \ref{fig:cloud-occurrence}b, c), we see that
the low-latitude SO zone displays a stronger peak of cloud occurrence near the
surface than the high-latitude SO zone, and this could be because higher
latitudes have less stable atmospheric profiles. The low- and high-latitude SO
zones show similar biases in models as in the general case, but ERA5 does not
overestimate the peak in the low-latitude SO zone (near-surface cloud
occurrence is still strongly underestimated).

@@

When subsetted by conditions of weak and strong stability (Fig.
\ref{fig:cloud-occurrence}f, g), as defined in Section
\ref{sec:cyclone-stability}, we see that in situations of strong stability
cloud occurrence peaks strongly at the surface in observations, compared to
situations of weak stability, where the peak is more diffuse between 0 and 1
km.  It is work mentioning that conditions of strong stability might be
associated with the formation of advection fog, such as in situation of warm
air advection from the north over colder sea surface, thus inducing fog
formation by cooling of the warm and humid air by the cold surface.  In
situations of strong stability, the models do a fairly good job, but
overestimate cloud occurrence at the peak below 1 km; above 1 km, they show
similar biases as in the general case.  In situations of weak stability, the
bias in ICON is very pronounced, with a much stronger peak at about 500 m, ERA5
underestimates cloud occurrence below 1 km (especially near the surface), and
MERRA-2 underestimates cloud occurrence even more strongly.

In all situations, even when the models overestimate cloud occurrence at some
altitudes, they always substantially underestimate the total cloud fraction.
ICON can be generally characterised as substantially overestimating cloud
occurrence below 1 km and underestimating above, underestimating the total
cloud fraction, and showing greatest biases in conditions of weak stability and
non-cyclonic conditions. It also shows a peak of cloud occurrence at higher
altitude than observations (500 m vs. near the surface), and correspondingly,
its LCL tends to be higher. MERRA-2 can be generally characterised as
underestimating cloud occurrence at nearly all altitudes as well as the total
cloud fraction, but mostly above and below 500 m (the peak at 500 m is
well-represented). MERRA-2 displays the largest errors relative to observations
in the low-latitude SO zone and in situations of weak stability. ERA5 can be
generally characterised as representing cloud occurrence correctly above about
1--1.5 km, overestimating below, but underestimating near-surface cloud
occurrence (0--500 m). The total cloud fraction is strongly underestimated in
all situations. ERA5 has a tendency towards underestimation in the low-latitude
SO zone and situations of weak stability; conversely, overestimating in the
high-latitude SO zone and conditions of strong stability.

\begin{figure}[p!]
\centering
\includegraphics[width=\textwidth]{img/cl_agg.pdf}
\caption{
Cloud occurrence by height calculated as the average of all voyages and
stations and lifting condensation level (LCL) distribution. The LCL is derived
from radiosonde profiles and equivalent model profiles, which were not
available for all voyages and times. The total cloud fraction (CF), average
shortwave (SW), and longwave (LW) and the relative frequency of occurrence
(RFO) are shown. The bands are the 16$^\mathrm{th}$--84$^\mathrm{th}$
percentile calculated from the set of all voyages and stations.
}
\label{fig:cloud-occurrence}
\end{figure}

@@

When analysed by subsets, observations in the cyclonic subset show the highest
cloud cover, with 8 oktas occurring on one half of such days (Fig.
\ref{fig:cloud-cover}d).  This sensitivity to cyclonic conditions is not
observed in ICON or the reanalyses. Interestingly, clear sky days (0 oktas)
also have a local maximum peaking at about 15\% in this subset.  When we
contrast the low- and high-latitude zones, we see that the high-latitude zone
tends to have greater cloud cover, peaking at 8 oktas (Fig.
\ref{fig:cloud-cover}c). The high-latitude zone also has almost no clear sky or
small cloud cover cases (0--4 oktas). ICON and the reanalyses represent this
characteristic of the distribution well for 0--3 oktas, but otherwise show
biases similar to the general case. One of the greatest biases is present in
ERA5 in the subset of weak stability, in which ERA5 peaks at 3 oktas, while the
observations peak at 7 oktas and show negligible cloud cover below 5 oktas.

@@

Relative humidity displays much larger biases. In all subsets, ICON is too
humid in the first 1 km by about 5\% but very accurate above, except for the
40--55°S zone and conditions of weak stability (Fig.
\ref{fig:potential-temperature}b, g),  where it is too dry between about 1 and
3 km. MERRA-2, on the other hand, is more humid than observations at all
altitudes and in all subsets, by up to about 20\% at 5 km. Even though the mean
near-surface relative humidity is similar to the observations (Fig.
\ref{potential-temperature}), the distribution in observation is more spread
out across both high and low values, and thus observations have a greater
prevalence of relative humidity close to 100\% and thus LCL located at the
surface (Fig. \ref{cloud-occurrence}a). In our calculations, LCL is a exclusive
function of near-surface temperature, near-surface relative humidity and
surface pressure.

@@

Compared to lidar observations, the daily cloud cover tends to be about 1 okta
lower in ICON and 2 oktas lower in the reanalyses.  Conditions of weak
stability are associated with some of the greatest biases, especially in ERA5.
The models also underestimate the cloud cover very strongly in cyclonic
conditions, which are very cloudy in the observations (8 oktas), but much less
so in the models.
